## Hi, decided to call the data set Food_Sec, have loaded the entire data set and would then subset it to the columns we are working on.
## There was one Raw_Data.csv file aswell that had done this already but that didn't have all the observations
getwd()

Food_Sec <-  data.frame(read.csv("DATS6101-Project1-2022Fall/dec21pub.csv"))

str(Food_Sec)
FS_Subset <- subset(Food_Sec, select = c("HRHHID",	"GTCO",	"GESTFIPS",	"HRNUMHOU",	"HEFAMINC",	"HESP1",	"PTDTRACE",	"PRCITSHP",	"PEMJNUM",	"PEHRUSL1",	"PEEDUCA", "HRFS12MD"))


## Installng tidyverse to use dplyr to rename the columns
install.packages('tidyverse')

library(dplyr)

head(FS_Subset)

FS_Subset <- FS_Subset %>% rename("Id" = "HRHHID", "Counties" = "GTCO", "States" = "GESTFIPS", "Family Size" = "HRNUMHOU",	"Household Income" = "HEFAMINC",	"SNAP" = "HESP1",	"Ethnicity" =	"PTDTRACE", "Citizenship status" = "PRCITSHP",	"Number of Jobs" = "PEMJNUM",	"Hours on Jobs" = "PEHRUSL1" , "Education Level" = "PEEDUCA" , "FoodSecurity score" = "HRFS12MD")

```{r, results='markup'}
head(FS_Subset)
str(FS_Subset)

```



## Converting the all the columns to factors as they are all ordinal(except the Id, but since it's categorical i'm converting it into a factor too)

FS_Subset[] <- lapply( FS_Subset, factor)


str(FS_Subset)


## Exploring the Counties indice

```{r, results='markup'}
summary(FS_Subset$Counties)
```


## The first row is the county number, check the documentation to figure out the county, and the row just beneath it is the count. I'll have to check what county 0 means, i have a feeling it might mean that no county was reported in the survey.
##Also not all counties are represented in this data. There are 3143 counties in the US in total, in this dataset we have 100 counties( including the 0 county).


## Installing ggplot, at then end we should just group all the library installation at the top

library(ggplot2)


ggplot(FS_Subset, aes(x=Counties)) + geom_bar()


