## Hi, decided to call the data set Food_Sec, have loaded the entire data set and would then subset it to the columns we are working on.
## There was one Raw_Data.csv file aswell that had done this already but that didn't have all the observations
getwd()
install.packages("epiR")
## Installng tidyverse to use dplyr to rename the columns
install.packages('tidyverse')

```{r}
Food_Sec <- data.frame(read.csv("dec21pub.csv"))
```

```{r}
library(dplyr)
library(ezids)
library(ggplot2)
library(epiR)
```

```{r, results='markup'}


FS_Subset <- subset(Food_Sec, HRINTSTA == 001 & HRSUPINT == 001 & HRFS12MD != -9)
FS_Subset <- subset(FS_Subset, select = c("HRHHID",	"GESTFIPS",	"HRNUMHOU",	"HEFAMINC",	"HESP1",	"PTDTRACE",	"PRCITSHP",	"PEMJNUM",	"PEHRUSL1",	"PEEDUCA", "HRFS12MD", "PRNMCHLD"))

FS_Subset <- FS_Subset %>% rename("Id" = "HRHHID", "States" = "GESTFIPS", "Family_Size" = "HRNUMHOU",	"Household_Income" = "HEFAMINC",	"SNAP" = "HESP1",	"Ethnicity" =	"PTDTRACE", "Citizenship_status" = "PRCITSHP",	"Number_of_Jobs" = "PEMJNUM",	"Hours_on_Jobs" = "PEHRUSL1" , "Education_Level" = "PEEDUCA" , "FoodSecurity_score" = "HRFS12MD")

## Converting the all the columns to factors as they are all ordinal(except the Id, but since it's categorical i'm converting it into a factor too)


FS_Subset[] <- lapply( FS_Subset, factor)

str(FS_Subset)
```


##Exploring the Food Security Score

```{r, results='markup'}
levels(FS_Subset$'FoodSecurity_score') <- c( "High Food Security", "Marginal Food Security", "Low Food Security", "Very Low Food Security")
summary(FS_Subset$'FoodSecurity_score')


FS_Subset$FS_Status <- FS_Subset$FoodSecurity_score

levels(FS_Subset$FS_Status) <- c( "Food Secure", "Food Secure", "Food Insecure", "Food Insecure")
```
  
##1 :High Food Security, 2 : Marginal Food Security , 3 : Low Food Security, 4  : Very Low Food Security ,-9 : No Response
## Children’s Food Security Scale variables are coded as “Not in Universe” (-1) if there were no children in the household. 
## Attachment 17, on page 288, contains details about the food security score. I have only given it a cursory glace, but we will have to read it in detail.




```{r, results='markup', }
ggplot(FS_Subset, aes(x= FoodSecurity_score,  y = ..prop.., group = 1)) + geom_bar( fill = "blue") +  geom_text(aes(label = round((..prop..), 2)), stat = "count", hjust = 0, colour = "black") + coord_flip()  + labs(x= " ", y= "Relative Frequency ") +  scale_y_continuous(limits = c(0, 1))
df <- FS_Subset %>% 
  group_by(FS_Status) %>% # Variable to be transformed
  count() %>% 
  ungroup() %>% 
  mutate(perc = `n` / sum(`n`)) %>% 
  arrange(perc) %>%
  mutate(labels = scales::percent(perc))

ggplot(df, aes(x = "", y = perc, fill = FS_Status)) +
  geom_col() +
    geom_text(aes(label = labels),
            position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y") + labs(x= " ", y= " ")



```

## Renaming levels of Ethnicity and Citizenship_Status
```{r, results='markup'}
levels(FS_Subset$'Ethnicity') <- c('White only', 'Black only', 'American Indian, Alaskan native only', 'Asian Only', 'Hawaiian', 'White-black', 'White-AI', 'White-Asian', 'White-HP', 'Black-AI', 'Black-Asian', 'Black-HP', 'AI-Asian', 'AI-HP', 'Asian-HP', 'W-B-AI', 'W-B-A', 'W-B-HP', 'W-AI-A', 'W-AI-HP', 'W-A-HP', 'B-AI-A', 'W-B-AL-A', 'W-AI-A-HP', 'Other 3 race combo', 'Other 4 and 5 race combo')
summary(FS_Subset$'Ethnicity', title = "PTDTRACE")

summary(FS_Subset$'SNAP', title = "HESP1")

levels(FS_Subset$'Citizenship_status') <- c('NATIVE, BORN IN THE UNITED STATES', 'NATIVE, BORN IN PUERTO RICO OR OTHER U.S. ISLAND AREAS', 'NATIVE, BORN ABROAD OF AMERICAN PARENT OR PARENTS', 'FOREIGN BORN, U.S. CITIZEN BY NATURALIZATION', 'FOREIGN BORN, NOT A CITIZEN OF THE UNITED STATES')
summary(FS_Subset$'Citizenship_status', title = "PRCITSHP")


```
## Plotting barcharts between all the levels of each factor and food security status

```{r, results='markup'}
table(FS_Subset$Ethnicity, FS_Subset$FoodSecurity_score)

for(i in levels(FS_Subset$Ethnicity)){
   
  print(ggplot(subset(FS_Subset, Ethnicity == i ) , aes(x= Ethnicity, color =FS_Status, fill = FS_Status)) + geom_bar( aes(y = (..count..)/sum(..count..)*100), position = "dodge") + labs(x= " ", y= "Percentage "))

}
```


```{r, results='markup'}

selected <- c('White only', 'Black only', 'American Indian, Alaskan native only', 'Asian Only', 'Hawaiian', 'White-black', 'White-AI', 'White-Asian', 'White-HP', 'Black-AI')

FS_ss <- FS_Subset[FS_Subset$Ethnicity %in% selected,]

FS_ss <- droplevels(FS_ss)


chi_test <- table(FS_ss$Ethnicity, FS_ss$FS_Status)
chi_test
chisq.test(chi_test)
fisher.test(FS_Subset$Ethnicity, FS_Subset$FoodSecurity_score, simulate.p.value=TRUE)

```


```{r, results='markup'}
table(FS_Subset$Citizenship_status, FS_Subset$FoodSecurity_score)

for(i in levels(FS_Subset$Citizenship_status)){
   
  print(ggplot(subset(FS_Subset, Citizenship_status == i ) , aes(x= Citizenship_status, color =FS_Status, fill = FS_Status)) + geom_bar( aes(y = (..count..)/sum(..count..)*100), position = "dodge")+ labs(x= " ", y= "Percentage "))
}




```

```{r, results='markup'}
chisq.test(FS_Subset$Citizenship_status, FS_Subset$FoodSecurity_score)
```


```{r, results='markup'}



## Only would be considering the levels 1 and 2. here -1, the one with most frequency, means that the response is not in universe. Ideally, I should substitute this but due to time constraints I won't be doing that. -3 and -2 mean refuse to respond and no response.


FS_SNAP <- subset(FS_Subset, FS_Subset$SNAP != -1)
FS_SNAP <- droplevels(FS_SNAP)

levels(FS_SNAP$SNAP) <- c('Refused', 'Dont Know', 'Uses SNAP', 'No SNAP')

table(FS_SNAP$SNAP, FS_SNAP$FoodSecurity_score)

ggplot(FS_SNAP, aes(x = SNAP, fill = FoodSecurity_score)) + geom_bar(position = "fill") + labs( x = " " , y = " ")
ggplot(FS_SNAP, aes(x = SNAP, fill = FS_Status)) + geom_bar(position = "fill") + labs( x = " " ,y = " ")
```

```{r, results='markup'}
chi_test_SNAP <- table(FS_SNAP$SNAP, FS_SNAP$FS_Status)
chi_test_SNAP

chisq.test(chi_test_SNAP)

```


```{r,results = 'markup'}
SNAP_Odds <- droplevels(subset(FS_SNAP, SNAP != "Refused" & SNAP != "Dont Know"))

TAB <- table(SNAP_Odds$SNAP, SNAP_Odds$FS_Status)
TAB

barplot(TAB, beside= T, legend = T)

epi.2by2(TAB, method = "cohort.count")
```

```{r data_ave, results='markup'}
food_ave <- subset(FS_Subset, select = c("Id",	"Number_of_Jobs",	"Hours_on_Jobs",	"Education_Level", "FoodSecurity_score"))
head(food_ave)
str(food_ave)
```

```{r factor, results='markup'}
food_ave$Number_of_Jobs = as.factor(food_ave$Number_of_Jobs)
food_ave$Education_Level = as.factor(food_ave$Education_Level)
food_ave$FoodSecurity_score = as.factor(food_ave$FoodSecurity_score)
food_ave$Hours_on_Jobs = as.numeric(food_ave$Hours_on_Jobs)
str(food_ave)
```

### Number of Jobs

```{r no_of_jobs, results='markup'}
plot(food_ave$Number_of_Jobs)
ggplot(food_ave) +
  geom_bar(aes(x = Number_of_Jobs, fill = FoodSecurity_score), position = "dodge") +
  xlab("Jobs") + ylab("Count")
```


```{r, results='markup'}
ggplot(subset(food_ave, Number_of_Jobs == -1)) +
  geom_bar(aes(x = Number_of_Jobs, fill = FoodSecurity_score), position = "dodge") +
  xlab("Jobs") + ylab("Count")

ggplot(subset(food_ave, Number_of_Jobs == 2)) +
  geom_bar(aes(x = Number_of_Jobs, fill = FoodSecurity_score), position = "dodge") +
  xlab("Jobs") + ylab("Count")

ggplot(subset(food_ave, Number_of_Jobs == 3)) +
  geom_bar(aes(x = Number_of_Jobs, fill = FoodSecurity_score), position = "dodge") +
  xlab("Jobs") + ylab("Count")

ggplot(subset(food_ave, Number_of_Jobs == 4)) +
  geom_bar(aes(x = Number_of_Jobs, fill = FoodSecurity_score), position = "dodge") +
  xlab("Jobs") + ylab("Count")
```

Here 
* 1 High Food Security
* 2 Marginal Food Security 
* 3 Low Food Security
* 4 Very Low Food Security
* -9 No Response


In the above graphs, people say that they have High Food Security irrespective of the number of jobs. But lets use Chi-Square test to see if they are really independent of each other

```{r, results='markup'}
contable <- table(food_ave$Number_of_Jobs, food_ave$FoodSecurity_score)
xkabledply(contable)

## Chi-Square Test 
chi_test <- chisq.test(contable, correct = F)
chi_test
```

The result gave warnings as the estimated value for some cells are very low.
From the test, we see that the P-value for the Chi-square test is 0.3871 which is greater than the default value 0.05. Hence we accept the null hypothesis and hence, Number of Jobs doesn't significantly affect the Food Security.


###Education Level

```{r, results='markup'}
edu_freq <- as.data.frame(table(food_ave$Education_Level))
edu_freq
```

```{r , results='markup'}
plot(food_ave$Education_Level)
ggplot(food_ave) +
  geom_bar(aes(x = Education_Level, fill = FoodSecurity_score), position = "dodge") +
  xlab("Jobs") + ylab("Count")
```
The response is understood as follows:

31 LESS THAN 1ST GRADE
32 1ST, 2ND, 3RD OR 4TH GRADE
33 5TH OR 6TH GRADE
34 7TH OR 8TH GRADE
35 9TH GRADE
36 10TH GRADE
37 11TH GRADE
38 12TH GRADE NO DIPLOMA
39 HIGH SCHOOL GRAD-DIPLOMA OR EQUIV (GED)
40 SOME COLLEGE BUT NO DEGREE
41 ASSOCIATE DEGREE-OCCUPATIONAL/VOCATIONAL
42 ASSOCIATE DEGREE-ACADEMIC PROGRAM
43 BACHELOR'S DEGREE (EX: BA, AB, BS)
44 MASTER'S DEGREE (EX: MA, MS, MEng, MEd, MSW)
45 PROFESSIONAL SCHOOL DEG (EX: MD, DDS, DVM)
46 DOCTORATE DEGREE (EX: PhD, EdD)

```{r, results='markup'}
ggplot(subset(food_ave, Education_Level == -1)) +
  geom_bar(aes(x = Education_Level, fill = FoodSecurity_score), position = "dodge") +
  xlab("Education") + ylab("Count")

ggplot(subset(food_ave, Education_Level == 31)) +
  geom_bar(aes(x = Education_Level, fill = FoodSecurity_score), position = "dodge") +
  xlab("Education") + ylab("Count")

ggplot(subset(food_ave, Education_Level == 32)) +
  geom_bar(aes(x = Education_Level, fill = FoodSecurity_score), position = "dodge") +
  xlab("Education") + ylab("Count")

ggplot(subset(food_ave, Education_Level == 33)) +
  geom_bar(aes(x = Education_Level, fill = FoodSecurity_score), position = "dodge") +
  xlab("Education") + ylab("Count")

ggplot(subset(food_ave, Education_Level == 34)) +
  geom_bar(aes(x = Education_Level, fill = FoodSecurity_score), position = "dodge") +
  xlab("Education") + ylab("Count")

ggplot(subset(food_ave, Education_Level == 35)) +
  geom_bar(aes(x = Education_Level, fill = FoodSecurity_score), position = "dodge") +
  xlab("Education") + ylab("Count")

ggplot(subset(food_ave, Education_Level == 36)) +
  geom_bar(aes(x = Education_Level, fill = FoodSecurity_score), position = "dodge") +
  xlab("Education") + ylab("Count")

ggplot(subset(food_ave, Education_Level == 37)) +
  geom_bar(aes(x = Education_Level, fill = FoodSecurity_score), position = "dodge") +
  xlab("Education") + ylab("Count")

ggplot(subset(food_ave, Education_Level == 38)) +
  geom_bar(aes(x = Education_Level, fill = FoodSecurity_score), position = "dodge") +
  xlab("Education") + ylab("Count")

ggplot(subset(food_ave, Education_Level == 39)) +
  geom_bar(aes(x = Education_Level, fill = FoodSecurity_score), position = "dodge") +
  xlab("Education") + ylab("Count")

ggplot(subset(food_ave, Education_Level == 40)) +
  geom_bar(aes(x = Education_Level, fill = FoodSecurity_score), position = "dodge") +
  xlab("Education") + ylab("Count")

ggplot(subset(food_ave, Education_Level == 41)) +
  geom_bar(aes(x = Education_Level, fill = FoodSecurity_score), position = "dodge") +
  xlab("Education") + ylab("Count")

ggplot(subset(food_ave, Education_Level == 42)) +
  geom_bar(aes(x = Education_Level, fill = FoodSecurity_score), position = "dodge") +
  xlab("Education") + ylab("Count")

ggplot(subset(food_ave, Education_Level == 43)) +
  geom_bar(aes(x = Education_Level, fill = FoodSecurity_score), position = "dodge") +
  xlab("Education") + ylab("Count")

ggplot(subset(food_ave, Education_Level == 44)) +
  geom_bar(aes(x = Education_Level, fill = FoodSecurity_score), position = "dodge") +
  xlab("Education") + ylab("Count")

ggplot(subset(food_ave, Education_Level == 45)) +
  geom_bar(aes(x = Education_Level, fill = FoodSecurity_score), position = "dodge") +
  xlab("Education") + ylab("Count")

ggplot(subset(food_ave, Education_Level == 46)) +
  geom_bar(aes(x = Education_Level, fill = FoodSecurity_score), position = "dodge") +
  xlab("Education") + ylab("Count")
```

```{r, results='markup'}
contable_edu <- table(food_ave$Education_Level, food_ave$FoodSecurity_score)
xkabledply(contable_edu)

## Chi-Square Test 
chi_test <- chisq.test(contable_edu, correct = F)
chi_test
```

`Education_Level` is having a significant effect on the Food Security of People

##Hours_on_Jobs

```{r , results='markup'}
plot(food_ave$Hours_on_Jobs)
ggplot(food_ave) +
  geom_bar(aes(x = Hours_on_Jobs, fill = FoodSecurity_score), position = "dodge") +
  xlab("Jobs") + ylab("Count")
```

```{r, results='markup'}
ggplot(food_ave, aes(Hours_on_Jobs, fill = FoodSecurity_score)) + 
        geom_histogram(binwidth=5)
ggplot(food_ave, aes(Hours_on_Jobs)) + 
        geom_freqpoly(binwidth=1)
ggplot(food_ave, aes(Hours_on_Jobs, color= FoodSecurity_score)) + 
        geom_freqpoly(binwidth=25)
ggplot(food_ave, aes(Hours_on_Jobs, fill = FoodSecurity_score)) + 
        geom_histogram(binwidth=5) + 
        facet_wrap(~FoodSecurity_score)
```

```{r, results='markup'}
vary_hours <- nrow(food_ave[food_ave$Hours_on_Jobs == -4,])
vary_hours
```

There are `r nrow(food_ave[food_ave$Hours_on_Jobs == -4,])` whose working hours vary. 

```{r, results='markup'}
summary(food_ave$Hours_on_Jobs)

getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
getmode(food_ave$Hours_on_Jobs)
```
```{r}
ggplot(food_ave, aes(x=FoodSecurity_score, y=Hours_on_Jobs)) + 
  geom_boxplot( colour=c("red","blue","violet","navy blue","black"), outlier.shape=8, outlier.size=4) +
  labs(title="Boxplot for Hours on Jobs", x="Control / Treatment", y = "Hours on Job")
```

```{r, results='markup', warning=FALSE}
h_anova <- aov(Hours_on_Jobs ~ FoodSecurity_score, data = food_ave)
h_tuskey <- TukeyHSD(h_anova)
summary(h_anova)
h_tuskey
```

2 - 1, 3 -1, 4 -1, 3 -2, 4 -2 have significant difference in there mean.
getmode(food_ave$Hours_on_Jobs)

```{r}
ggplot(food_ave, aes(x=FoodSecurity_score, y=Hours_on_Jobs)) + 
  geom_boxplot( colour=c("red","blue","violet","navy blue","black"), outlier.shape=8, outlier.size=4) +
  labs(title="Boxplot for Hours on Jobs", x="Control / Treatment", y = "Hours on Job")
```


There are `r nrow(food_ave[food_ave$Hours_on_Jobs == -4,])` whose working hours vary. 
summary(food_ave$Hours_on_Jobs)

>>>>>>> 8657d5728a00f036f19d3ba04f8e0d67a4b3431f


-***States, Family size, and Household Income***

```{r, results='markup'}
levels(FS_Subset$'States') <- c('AL', 'AK', 'AZ', 'AR', 'CA', 'CO', 'CT', 'DE', 'DC', 'FL', 'GA', 'HI', 'ID', 'IL', 'IN', 'IA', 'KS', 'KY', 'LA', 'ME', 'MD', 'MA', 'MI', 'MN', 'MS', 'MO', 'MT', 'NE', 'NV', 'NH', 'NJ', 'NM', 'NY', 'NC', 'ND', 'OH', 'OK', 'OR', 'PA', 'RI', 'SC', 'SD', 'TN', 'TX', 'UT', 'VT', 'VA', 'WA', 'WV', 'WI', 'WY' )
summary(FS_Subset$'States', title = "GESTFIPS")

```

California has the highest number of respondents (6975), whereas Maine has the smallest number of respondents (564). 
In order to compare, I'm choosing states which has similar number of respondents. 
Alabama 1207 and Washington DC 1207, Florida 2738 and New York 2580, IL 2052 and PA 1928. 

## Plotting barcharts between all the levels of state and food security status

```{r, results='markup'}
for(i in levels(FS_Subset$States)){
   
  print(ggplot(subset(FS_Subset, States == i ) , aes(x= States, color =FoodSecurity_score, fill = FoodSecurity_score)) + geom_bar( aes(y = (..count..)/sum(..count..)*100), position = "dodge"))
}
```


```{r data_una, results='markup'}
library(ggplot2)
#qqplot(food_una$States, food_una$FoodSecurity_score, plot.it = TRUE,  xlab="States", ylab="Food security Score",  main="QQ Plot for Food security vs States")
hist(x=States, )
```
<<<<<<< HEAD



```{r data_una, results='markup'}
#food_una <- subset(FS_Subset, select = c("HRHHID",	"GESTFIPS", "HRNUMHOU",	"HEFAMINC", "HRFS12MD"))
food_una <- subset(FS_Subset, select = c("Id", "States","Family_Size","Household_Income","FoodSecurity_score", "PRNMCHLD"))
head(food_una)

```

Reference to household income: 
1 LESS THAN $5,000  
2 5,000 TO 7,499  
3 7,500 TO 9,999  
4 10,000 TO 12,499  
5 12,500 TO 14,999  
6 15,000 TO 19,999  
7 20,000 TO 24,999  
8 25,000 TO 29,999  
9 30,000 TO 34,999  
10 35,000 TO 39,999  
11 40,000 TO 49,999  
12 50,000 TO 59,999  
13 60,000 TO 74,999  
14 75,000 TO 99,999  
15 100,000 TO 149,999  
16 150,000 OR MORE  

```{r}
income_t <- table(FS_Subset$Household_Income, FS_Subset$FoodSecurity_score)
chisq.test(income_t)
```
We can say that Household income is affecting food insecurity. 

```{r}
family_t <- table(FS_Subset$Family_Size, FS_Subset$FoodSecurity_score)
chisq.test(family_t)
```
We can say that Family size is affecting food insecurity. 

```{r data_una, results='markup'}

boxplot(FoodSecurity_score ~ Family_Size, data = food_una, col=c("blue", "red", "green", "yellow", "purple", "orange"), main="Family size - Food security")
boxplot(FoodSecurity_score ~ Household_Income, data = food_una, col=c("blue", "red", "green", "yellow", "purple", "orange"), main="Household Income - Food security")
boxplot(FoodSecurity_score ~ States, data = food_una, col=c("blue", "red", "green", "yellow", "purple", "orange"))

```
As you can see from the boxplot, whenever family size bigger (more than 6 people), food insecurity is high. 
Also, household income has direct effect on food security. When household income is higher than 40k, food security score is low. 

```{r data_una, results='markup'}
library(ggplot2)
qqplot(food_una$Household_Income, food_una$FoodSecurity_score, plot.it = TRUE,  xlab="Household income", ylab="Food security Score",  col=food_una$Household_Income, main="QQ Plot for Food security vs Household income")

```

```{r data_una, results='markup'}
summary(FS_Subset$Family_Size)

```


```{r data_una, results='markup'}
library(ggplot2)
qqplot(food_una$Family_Size, food_una$FoodSecurity_score, plot.it = TRUE,  xlab="Family size", ylab="Food security Score",  col=food_una$Family_Size, main="QQ Plot for Food security vs Family")

```
```{r data_una, results='markup'}

ggplot(data=food_una, mapping = aes(x = Family_Size, y = FoodSecurity_score)) + 
  geom_point(size=5)+aes(color = Family_Size) +
  theme_bw()

```

```{r data_una, results='markup'}

ggplot(data=food_una, mapping = aes(x = Family_Size, y = FoodSecurity_score)) + 
  geom_point(size=5)+aes(color = Household_Income) +
  theme_bw()

```


