* Hi, this is the rmd file for project 2. We will be doing logistic regression to acertain the likelihood of food insecurity in permutations of social factors



```{r, echo=FALSE, cache=TRUE }
Food_Sec <- data.frame(read.csv("dec21pub.csv"))
```



```{r, echo=FALSE, results='hide', message=FALSE}
library(dplyr)
library(ezids)
library(ggplot2)
library(epiR)
```

* Lets start with the variables we selected for the first project. We should replace some variables I think, but we'll do it along the way. I am just copying code from the first project for now. so we can get a base to build upon.

```{r, results='markup', echo=FALSE}


FS_Subset <- subset(Food_Sec, HRINTSTA == 001 & HRSUPINT == 001 & HRFS12MD != -9)
FS_Subset <- subset(FS_Subset, select = c("HRHHID",	"GESTFIPS",	"HRNUMHOU",	"HEFAMINC",	"HESP1",	"PTDTRACE",	"PRCITSHP",	"PEMJNUM",	"PEHRUSL1",	"PEEDUCA", "HRFS12MD", "PRNMCHLD"))

FS_Subset <- FS_Subset %>% rename("Id" = "HRHHID", "States" = "GESTFIPS", "Family_Size" = "HRNUMHOU",	"Household_Income" = "HEFAMINC",	"SNAP" = "HESP1",	"Ethnicity" =	"PTDTRACE", "Citizenship_status" = "PRCITSHP",	"Number_of_Jobs" = "PEMJNUM",	"Hours_on_Jobs" = "PEHRUSL1" , "Education_Level" = "PEEDUCA" , "FoodSecurity_score" = "HRFS12MD")

## Converting the all the columns to factors as they are all ordinal(except the Id, but since it's categorical i'm converting it into a factor too)


FS_Subset[] <- lapply( FS_Subset, factor)

str(FS_Subset)
```
<<<<<<< Updated upstream
* Since we will be doing logistic regression, we have to reduce our response to two level only. If we complete this project before the deadline we can look into multinomial regression aswell.
=======
* Since we will be doing logistic regression, we have to reduce our response to two level only. If we complete this project before the deadline we can look into multinational regression aswell.
>>>>>>> Stashed changes

```{r, results='markup', echo=FALSE, results='hide'}
levels(FS_Subset$'FoodSecurity_score') <- c( "High Food Security", "Marginal Food Security", "Low Food Security", "Very Low Food Security")
summary(FS_Subset$'FoodSecurity_score')


FS_Subset$FS_Status <- FS_Subset$FoodSecurity_score

levels(FS_Subset$FS_Status) <- c( "Food Secure", "Food Secure", "Food Insecure", "Food Insecure")
```